---
title: "Regular Expressions and Data Manipulation"
author: "Kyle Samson"
date: "2025-01-08"
output: html_document
---

# Introduction
This document demonstrates how to use R and regular expressions to manipulate data in `measurements.csv`. The steps include:

- Importing the dataset
- Using `dplyr` and regex to modify the `Organism` column
- Reordering columns and saving the modified dataset
- Committing and pushing changes to GitHub

# Load Required Libraries
```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
```

# Step 2: Import Data
```{r}
# Read the CSV file into an object called MData
MData <- read.csv("measurements.csv")

# Display first few rows to confirm import
head(MData)
```

# Step 3: Modify the Organism Column Using Regex
```{r}
# Use regex with dplyr to create a new column "Sp" that shortens genus names
MData <- MData %>%
  mutate(Sp = sub("^(\\w)\\w+ (\\w+)", "\\1. \\2", Organism))

# Check the first few rows to ensure transformation worked
head(MData)
```

# Step 4: Reorder Columns
```{r}
# Create a new dataset with 'Sp' replacing 'Organism' in the first column
MData_new <- MData %>%
  select(Sp, everything(), -Organism)  # Move 'Sp' to the front and remove 'Organism'

# Display first few rows to verify column order
head(MData_new)
```

# Step 5: Save Modified Dataset
```{r}
# Save the new dataset as MeasShort.csv
write.csv(MData_new, "MeasShort.csv", row.names = FALSE)
```


# Conclusion
This report demonstrates how to use `dplyr` and regular expressions in R to clean and process data efficiently. The modified dataset has been saved as `MeasShort.csv`, and all changes have been pushed to GitHub.

